\ Legacy Data Confounds Genomics Studies

### 2019-04-05 legacy.tex
MBE formatting

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ELIFE ARTICLE TEMPLATE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% PREAMBLE 
\documentclass[9pt,lineno]{elife}

\usepackage[export]{adjustbox}
\usepackage{lscape}
\usepackage{afterpage}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}

\newcommand{\sgcomment}[1]{\textcolor{blue}{SG: #1}}
\newcommand{\luke}[1]{\textcolor{blue}{Luke: #1}}
\newcommand{\todo}[1]{\textcolor{blue}{*#1*}}
\newcommand{\alex}[1]{\textcolor{red}{Alex: #1}}
\newcommand{\logit}{\operatorname{logit}}
%\newcommand{\qq}{m}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE SETUP
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Legacy Data Confounds Genomics Studies}

\author[1,2]{Luke Anderson-Trocm\'e}
\author[1,2]{Rick Farouni}
\author[1,2]{Mathieu Bourgey}
\author[3]{Fumihiko Matsuda}
\author[3]{Yoichiro Kamatani}
\author[1,2]{Simon Gravel}

\affil[1]{Department of Human Genetics, McGill University, Montreal, QC H3A 0G1, Canada}
\affil[2]{McGill University and Genome Quebec Innovation Centre, Montreal, QC H3A 0G1, Canada}
\affil[3]{Center for Genomic Medicine, Graduate School of Medicine, Kyoto University, Kyoto 606-8501, Japan}
\corr{simon.gravel@mcgill.ca}{SG}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE START
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\maketitle
\begin{abstract}
Recent reports have identified differences in the mutational spectra across human populations. While some of these reports have been replicated in other cohorts, most have been reported only in the 1000 Genomes Project (1kGP) data. While investigating an intriguing putative population stratification within the Japanese population, we identified a previously unreported batch effect leading to spurious mutation calls in the 1kGP data and to the apparent population stratification. Because the 1kGP data is used extensively, we find that the batch effects also lead to incorrect imputation by leading imputation servers and suspicious GWAS associations. Lower-quality data from the early phases of the 1kGP thus contaminates modern studies in hidden ways, and a community effort may be required to remove or upgrade such legacy sequencing data from reference databases. 
\end{abstract}

@@

\section{Methods}
\subsection{Metadata}
The metadata used in this analysis was compiled from each of the index files from the 1000 Genomes file system. 
Average quality of mapped bases $Q$ per sample was obtained from the BAS files associated with each alignment file. 
Each BAS file has metadata regarding each sequencing event for each sample. 
If a sample was sequenced more than once, we took the average of the each $Q$ score from each sequencing instance. 
The submission dates and sequencing centres for each sample in the analysis was available in the sequence index files \ref{ProcessPlotData.Rmd}.

@@

\bibliography{Legacy.bib}

>>>>

% PREAMBLE 
%\documentclass[9pt,lineno]{elife}
\documentclass[custompaper]{MBE}%
%\DeclareOption{a4paper}
\usepackage{url}
\usepackage{xcolor}
\usepackage{subcaption}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}

\newcommand{\sgcomment}[1]{\textcolor{blue}{SG: #1}}
\newcommand{\luke}[1]{\textcolor{blue}{Luke: #1}}
\newcommand{\logit}{\operatorname{logit}}

%\jshort{mst}

%\volname{}

%\jvolume{0}

%\jvol{}

%\jissue{0}

%\pubyear{2019}

\mstype{Article}

%\artid{012}

%\access{Advance Access publication March 3, 2013}

@@

\keyword{1000 Genomes Project, Batch Effect, Population Structure, GWAS, GCAT}

\title{\ Legacy Data Confounds Genomics Studies}

\author[Anderson-Trocm\'e et al.]{\ Luke \surname{Anderson-Trocm\'e},$^{\ast,1,2}$ Rick Farouni,$^{1,2}$ Mathieu Bourgey,$^{1,2}$ Fumihiko Matsuda,$^{3}$, Yoichiro Kamatani$^{3}$, and Simon Gravel$^{1,2}$,}

\address{$^{1}$Department of Human Genetics, McGill University, Montreal, QC H3A 0G1, Canada\\
$^{2}$McGill University and Genome Quebec Innovation Centre, Montreal, QC H3A 0G1, Canada\\ 
$^{3}$Center for Genomic Medicine, Graduate School of Medicine, Kyoto University, Kyoto 606-8501, Japan}

\coresp{simon.gravel@mcgill.ca}

\abstract{Recent reports have identified differences in the mutational spectra across human populations. While some of these reports have been replicated in other cohorts, most have been reported only in the 1000 Genomes Project (1kGP) data. While investigating an intriguing putative population stratification within the Japanese population, we identified a previously unreported batch effect leading to spurious mutation calls in the 1kGP data and to the apparent population stratification. Because the 1kGP data is used extensively, we find that the batch effects also lead to incorrect imputation by leading imputation servers and suspicious GWAS associations. Lower-quality data from the early phases of the 1kGP thus contaminates modern studies in hidden ways, and a community effort may be required to remove or upgrade such legacy sequencing data from reference databases. }

\maketitle

@@

\newpage{}
\section{Methods}
\subsection{Metadata}
The metadata used in this analysis was compiled from each of the index files from the 1000 Genomes file system. 
Average quality of mapped bases $Q$ per sample was obtained from the BAS files associated with each alignment file. 
Each BAS file has metadata regarding each sequencing event for each sample. 
If a sample was sequenced more than once, we took the average of the each $Q$ score from each sequencing instance. 
The submission dates and sequencing centres for each sample in the analysis was available in the sequence index files \ref{ProcessPlotData.Rmd}.

@@

\bibliographystyle{natbib}
\bibliography{Legacy.bib}
